Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-12-01 paper/main.tex
reshuffled related-work section

\section{Related Work}
This work aims to develop a model for describing visualization transformations that can serve as guidance for how to architecture a general purpose visualization library. We define a general purpose visualization library as one that provides non domain specific building block components\cite{wongsuphasawatNavigatingWideWorld2021} for building visualizations, for example functions for converting data to color or encoding data as dots. In this section, we describe how visualization libraries attempt this goal and discuss work that formally describes what properties of data should be preserved in a visualization. We restrict the properties of data that should be preserved to

\begin{LaTeXdescription}
  \item [continuity] how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
  \item [equivariance] functions on data that have an equivalent effect on the graphical representation, e.g. rotating a matrix has a matching rotation of the image, translating the points on a line has a matching visual shift in the line plot
\end{LaTeXdescription}

\subsection{Continuity}
\begin{figure}[!h]
  \includegraphics[width=\columnwidth]{k_different_types.png}
  \caption{Continuity is how elements in a data set are connected to each other, which is distinct from how the data is structured. The rows in (a) are discrete, therefore they have discrete continuity as illustrated by the discrete dots. The gaussian in (b) is a 1D continuous function, therefore the continuity of the elements of the gaussian can be represented as a line on an interval (0,1). In (c), every element of the globe is connected to its nearest neighbors, which yields a 2D continuous continuity as illustrated by the square.}
  \label{fig:related-work:continuity}
\end{figure}
Continuity describes elements in a data set are organized; this concept is termed topological properties by Wilkinson\cite{wilkinsonGrammarGraphics2005}. Wilkinson provides the examples of values that are isolated from each other, and therefore discrete, and values lying on a continuum or in a compact region. For example, in \autoref{fig:related-work:continuity}, each station record in the table is independent of the others; therefore, the continuity of the table is discrete. The data provided by the gaussian are points sampled along the curve, therefore the continuity of the points on the line is 1D continuous. Every point on the globe is connected to its 6 nearest cardinal neighboring points (NW, N, NE, E, SE, S, SW, W). We propose that a robust model of continuity provides a way to develop library components that can work on the data in small pieces in a manner where the overall continuity of the data is preserved in the visual transformation.

@@

@@

A fiber bundle has many sections, and the set of all sections over an open set $openset \subseteq \dtotal$
\begin{equation}
\label{eq:related-work:continuity:sections}
\cgamma{\opensetc}{\dtotalc} \coloneqq \{\dsectionc^{i}: \opensetc\rightarrow \dtotalc\}|_{i\in I}
\end{equation}
 is the set of all data sets with a shared fiber \dfiber\ and continuity \dbase. The set of all sections defined over the total base space is denoted $\Gamma(\dbase, \dtotal)$.

>>>>

\section{Related Work}
This work aims to develop a model for describing visualization transformations that can serve as guidance for how to architecture a general purpose visualization library. We define a general purpose visualization library as one that provides non domain specific building block components\cite{wongsuphasawatNavigatingWideWorld2021} for building visualizations, for example functions for converting data to color or encoding data as dots. We propose that visualization library components should preserve continuity and equivariance, which we define as

\begin{LaTeXdescription}
  \item [continuity] how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
  \item [equivariance] data transformations that have an equivalent effect on the graphical representation, e.g. rotating a matrix has a matching rotation of the image, translating the points on a line has a matching visual shift in the line plot
\end{LaTeXdescription}

In \autoref{sec:related-work:continuity} we discuss how data continuity influences visualization library architecture and in \autoref{sec:related-work:equivariance} we summarize how visualization researchers have approached equivariance.

\subsection{Continuity}
\label{sec:related-work:continuity}
\begin{figure}[!h]
  \includegraphics[width=\columnwidth]{k_different_types.png}
  \caption{Continuity is how elements in a data set are connected to each other, which is distinct from how the data is structured. The rows in (a) are discrete, therefore they have discrete continuity as illustrated by the discrete dots. The gaussian in (b) is a 1D continuous function, therefore the continuity of the elements of the gaussian can be represented as a line on an interval (0,1). In (c), every element of the globe is connected to its nearest neighbors, which yields a 2D continuous continuity as illustrated by the square.}
  \label{fig:related-work:continuity}
\end{figure}
Continuity describes elements in a data set are organized; this concept is termed topological properties by Wilkinson\cite{wilkinsonGrammarGraphics2005}. Wilkinson provides the examples of values that are isolated from each other, and therefore discrete, and values lying on a continuum or in a compact region. For example, in \autoref{fig:related-work:continuity}, each station record in the table is independent of the others; therefore, the continuity of the table is discrete. The data provided by the gaussian are points sampled along the curve, therefore the continuity of the points on the line is 1D continuous. Every point on the globe is connected to its 6 nearest cardinal neighboring points (NW, N, NE, E, SE, S, SW, W). We propose that a robust model of continuity provides a way to develop library components that can work on the data in small pieces in a manner where the overall continuity of the data is preserved in the visual transformation.

@@

\section{Algebraic Topology And Category Theory}

@@

A fiber bundle has many sections, and the set of all sections over an open set $\openset \subseteq \dtotal$
\begin{equation}
  \label{eq:related-work:continuity:sections}
  \begin{split}
  \cgamma{\opensetc}{\dtotalc} \coloneqq & \{\dsectionc^{0} \cdots \dsectionc^{i} \cdots \dsectionc^{n}\} 
  & \dsectionc: \opensetc\rightarrow \dtotalc
  \end{split}
\end{equation}
 is the set of all data sets with a shared fiber \dfiber\ and continuity \dbase. The set of all sections defined over the total base space is denoted $\Gamma(\dbase, \dtotal)$.
